---
title: "CONN_MPC_Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CONN_MPC_Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Basic distance and connectivity functions in the Reconnect R-package
Demonstration of application of functions described in the CONN_functions.R and MPC_functions.R files.


```{r setup}
#########################################################
### clean space
rm(list=ls(all=TRUE))
#########################################################
## set directories
## make sure you have loaded the "MSC_analysis.Rproj" correctly
if(!exists('sourcedir')){ sourcedir <- getwd() }; if(!exists('resdir')) { resdir <- getwd() }
##########################################################
## libraries
library(sf)               # new spatial vector processing package
library(raster)           # old (but still useful) spatial raster processing
library(stars)            # spatial processing (conversion between sf and raster possible, spatiotemporal datacube analyses)
library(terra)
library(igraph)
library(Reconnect)
## some colors...
JOcols  = c("cyan3","olivedrab3","olivedrab4")
JOcols1 = colorRampPalette(c("navy","lightblue","grey","mediumorchid","cyan3","cyan4","tan","olivedrab3","olivedrab2","olivedrab4"))
```

## Create a simple landscape
e.g. from the simple circle simulation in the NLMC_functions.R file

```{r makelandscape}

######################################################################
## set seed for reproducibility
set.seed(15)
### create a name of the simulation
## initialize parameters
nc1     = 100                                       ## define number of cells of landscape # 500 cells for dimx and dimy is reasonably good!!
hfr1    = c(0.15)                                   ## define fraction of habitat in landscape, in combination with npatches, this defines patch area
npatches= c(3) ## define number of patches (maximum nr should not be more than 10 times the number of cells..) in combination with hfrs, this defines patch area
sdpa1   = 0                                         ## define sd of patch area
cf1     = 1                                         ## clumping factor: 1 = non-overlapping patches, >1: clumped & overlapping patches
remedge1= TRUE                                      ## remove the edge of the landscape? - i.e. patches cannot go partially outside landscape

### prepare plot
par(mfrow=c(1,1))
### make a simple circle simulation
for(npatch1 in npatches){
  res = simpcirc(dimx=nc1,dimy=nc1,hfr=hfr1,npatch=npatch1,sdpa=sdpa1,cf=cf1,form="circle",return="all",remedge=remedge1)
  ### unify new shape
  sps1 = sf::st_union(res$sps,by_feature=FALSE) %>% st_cast("POLYGON") %>% st_sf # st_cast(sps1,"POLYGON")
  ## total new area
  #tna   = sum(res$rast[res$rast==1])
  tna    = sum(as.numeric(sf::st_area(sps1))) # new area
  ## total nr of new patches
  tnpatch = length(sps1$geometry)
  ##plot result
  plot(res$rast,main=sprintf("total area (ha) =%0.0f habitat fraction = %0.02f \n nr of patches = %0.0f clumping factor = %0.02f",tna,hfr1,npatch1,cf1),cex.main=0.8)
  plot(sps1$geometry,add=TRUE,border="darkgreen")
}

### from this simulation, you can derive habitat rasters and shapefiles
## habitat shapefile
lcshp = res$sps
plot(lcshp,main="habitat shapefile")
## for plotting later
ccoord = st_coordinates(st_centroid(lcshp))
labels = lcshp$ID

## habitat raster
lcr = res$rast
#print(lcr)
plot(lcr,main="habitat landscape")

## resitance raster
lres = res$rast
## set all to 1 for the moment
lres[] = 1
## add a barrier whith resistance 10: cost distances should be 20 meters longer!
lres[5001:5099] = 20
#print(lres)
plot(lres,main="resistance landscape")
plot(lcshp$geometry,add=TRUE)
text(ccoord,labels=labels)

## set roi to NULL to get the full extent
# bbox = sf::st_bbox(lcr)
# bbox[c(3,4)] = bbox[c(1,2)]+0.8*(bbox[c(3,4)]-bbox[c(1,2)])
# roi  = sf::st_sf(ID= 1, geometry=sf::st_as_sfc(bbox))
# sf::st_crs(roi) = sf::st_crs(lcr)
# print(roi)
# plot(roi$geometry,add=TRUE)
roi = NULL

```


## Apply some distance functions
e.g. by using functions in the CONN_functions.R file

```{r distance functions}

### get euclidean distance among habitat patches!!
eucpm = eucdist_pm(x=lcshp,id="ID", dist_type = "centroid")
print("euclidean distance, dist_type=centroid")
print(eucpm)

eucpm = eucdist_pm(x=lcshp,id="ID", dist_type = "edge")
print("euclidean distance, dist_type=edge")
print(eucpm)

### get cost distance among habitat patches!!
costpm = costdist_pm(x=lcshp,id="ID", resi=lres, dist_type = "centroid")
print("cost distance, dist_type=centroid")
print(costpm)

costpm = costdist_pm(x=lcshp,id="ID", resi=lres, dist_type = "edge")
print("cost distance, dist_type=edge")
print(costpm)

### get euclidean distance for each pixel in landscape!
euclm  = eucdist_lm(x=lcr,exclude=NULL,dist_type = "edge")
## create a null raster to fill in values
lcr0   = lcr
lcr0[] = euclm[,1]
plot(lcr0, main="euclidean shortest \n distance to habitat")
plot(lcshp$geometry,add=TRUE)
text(ccoord,labels=labels)

### get euclidean distance from habitat centers to edge of habitat!
euchab = eucdist_hab(x=lcr,exclude=NULL)
lcr0   = lcr
lcr0[] = euchab[,1]
plot(lcr0, main="euclidean shortest \n distance to habitat edge")
plot(lcshp$geometry,add=TRUE)
text(ccoord,labels=labels)

### get cost distance at landscape scale!!
costlm = costdist_lm(x=lcr,resi=lres,scale=1,maxiter=100,dist_type="edge",cost_type ="least-cost")
lcr0   = lcr
lcr0[] = costlm[,1]
plot(lcr0, main="resistance shortest \n distance to habitat")
plot(lcshp$geometry,add=TRUE)
text(ccoord,labels=labels)

```


## Apply some connectivity functions
e.g. by using functions in the CONN_functions.R & MPC_functions.R files

```{r connectivity functions}

## graph centrality functions (betweenness centrality, closeness centrality and node degree from igraph)
centr = centr_igraph(pa=eucpm$pa,mdist=eucpm$mdist,alpha=40,dispfop="negex",weighted=TRUE,cutoffpr=0.01,MST=FALSE)
plot(lres,legend=FALSE,main="closeness centrality \n  euclidean distance")
plot(lcshp$geometry,col=JOcols[order(centr$CC,decreasing=FALSE)],main="closeness centrality euclidean distance",add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(centr$CC,decreasing=FALSE),2),bty="n")

centr = centr_igraph(pa=costpm$pa,mdist=costpm$mdist,alpha=40,dispfop="negex",weighted=TRUE,cutoffpr=0.01,MST=FALSE)
plot(lres,legend=FALSE,main="closeness centrality \n cost distance")
plot(lcshp$geometry,col=JOcols[order(centr$CC,decreasing=FALSE)],main="closeness centrality cost distance",add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(centr$CC,decreasing=FALSE),2),bty="n")

#####
## Equivalent connected area, Probability of connectivity, ProtConn (i.e. ECA_AL) functions...
## numerator of probability of connectivity (needs to be divided by the area of the landscape to yield Probability of Connectivity,  PC = pcnum/(AL)^2)
# pcnum = PCnum(pa=eucpm$pa,mdist=eucpm$mdist,alpha=40,dispfop="negex",savememory=FALSE)
# ECA_pc(pa=eucpm$pa,mdist=eucpm$mdist,alpha=40,dispfop="negex",savememory=FALSE)
# ECA_pc(pcnum=pcnum)
## wrapper function for all metrics at once:
#PCECA_fun(x=lcr,resi=lres,cost_type="least-cost")
## in order to speed up calculations, directly take distance matrices created above, but then you need to provide area of AL
AL    = sf::st_area(st_as_sfc(st_bbox(lcr)))
pceca = PCECA_fun(pa=costpm$pa,mdist=costpm$mdist,AL=AL)
print(pceca)

######
## patch importance for function dI - there is a dI_fun and a dI_fun_alpha, where dispersal capacities can be indicated!
## this function can be slow if there are many patches (more than 10000...)
## define a connectivity function
confun = "PCnum"
## define an input this function can handle (it can also just be x = lcr)
x  = list(pa=costpm$pa,mdist=costpm$mdist)
## importantly, the user needs to specify "k", i.e. how list objects should be subsetted;
expression = "pa=x[[1]][-k],mdist=x[[2]][-k,-k],alpha=alpha"
pid = colnames(costpm$mdist)
resnr = 1
alpha0 = 40
dIPCnum = dI_fun_alpha(fun=confun,expr=expression,x=x,pid=pid,resnr=resnr,alpha=alpha0)
## plot for overview
plot(lres,legend=FALSE,main="Patch importance for PCnum, \n  based on cost distance")
plot(lcshp$geometry,col=JOcols[order(dIPCnum,decreasing=FALSE)],add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(dIPCnum,decreasing=FALSE),2),bty="n")

### try it with more difficult function - ECA and PC!
confun = "PCECA_fun"
x  = list(pa=costpm$pa,mdist=costpm$mdist)
expression = "pa=x[[1]][-k],mdist=x[[2]][-k,-k],alpha=alpha,AL=10000"
pid     = colnames(x[[2]])
resnr   = c(2,3)
alpha0  = 40
dIPCECA = dI_fun_alpha(fun=confun,expr=expression,x=x,pid=pid,resnr=resnr,alpha=alpha0)
print(dIPCECA)
dIECA   = dIPCECA[,2]
## plot for overview
plot(lres,legend=FALSE,main="Patch importance for ECA, \n based on cost distance")
plot(lcshp$geometry,col=JOcols[order(dIECA,decreasing=FALSE)],add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(dIECA,decreasing=FALSE),2),bty="n")


####
## inverse cumulative resistance between x number of pairs in habitat network invCR
print("inverse cumulative resitance \n among all habitat patches")
print(invCR_p(pa=costpm$pa,mdist=costpm$mdist,pnr=1))

### patch importance is possible here as well!
confun = "invCR_p"
x  = list(pa=costpm$pa,mdist=costpm$mdist)
expression = "pa=x[[1]][-k],mdist=x[[2]][-k,-k],pnr=1"
pid   = colnames(x[[2]])
resnr = c(1)
dIinvCR_p = dI_fun(fun=confun,expr=expression,x=x,pid=pid,resnr=resnr)
## this has some negative value...
dIinvCR_p = dIinvCR_p + abs(min(dIinvCR_p)) + 1
## plot for overview
plot(lres,legend=FALSE,main="Patch importance for inverse \n cumulative resistance, based on cost distance")
plot(lcshp$geometry,col=JOcols[order(dIinvCR_p,decreasing=FALSE)],add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(dIinvCR_p,decreasing=FALSE),2),bty="n")

####
## Metapopulation metrics!
mpcres = MPC_fun(pa=costpm$pa,mdist=costpm$mdist,alpha=40)
## plot for overview
plot(lres,legend=FALSE,main="Patch importance for metapop. capacity, \n based on cost distance")
plot(lcshp$geometry,col=JOcols[order(mpcres$pimport*100,decreasing=FALSE)],add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(mpcres$pimport*100,decreasing=FALSE),2),bty="n")

####
## calculate a series of Metapopulation Capacity (MPC) - based metrics
## define region of interest
roi     = st_sf(id=1,geometry=st_as_sfc(st_bbox(lcr)))
## define baseline MPC value (e.g. if all the landscape would be habitat)
MPCbasl =  MPC_fun(x=roi)$mpc
###
mpcres = MPCser(pa=costpm$pa,mdist=costpm$mdist,alpha=40,roi=roi,MPCbasl=MPCbasl)
## plot for overview
plot(lres,legend=FALSE,main="Patch importance for metapop. capacity, \n based on cost distance")
plot(lcshp$geometry,col=JOcols[order(mpcres$MPCev*100,decreasing=FALSE)],add=TRUE)
text(ccoord,labels=labels)
legend("topleft",fill=JOcols,legend=round(sort(mpcres$MPCev*100,decreasing=FALSE),2),bty="n")


## some further connectivity metrics..
print("some further connectivity metrics...")

## inverse resistance / distance of landscape...
print(invCR_l(x=lcr,resi=lres,cost_type = "least-cost",dist_type="edge"))

## some general Landscape statistics metrics (patch area, number of patches, nearest distances...)
print(LandStats(pa=costpm$pa,mdist=costpm$mdist))

## some general Fragstats metrics (clumpiness index, cohesion index, mesh size, percent of like adjacencies...)
print(FragStats(x=lcr))

```
