---
title: "NLMC_Demo"
author: "Jacqueline Oehri"
date: "2023-08-01"
output:
  html_document:
   # theme: journal
   # highlight: tango
   # toc: true
   # toc_float: true
   # number_sections: true
    code_folding: hide
    code_download: true
#runtime: shiny
---


```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Neutral Landscape Model (NLMC) computation

The Reconnect R-package offers 2 ways to generate simulated landscapes where the amount and fragmentation of habitat can be adjusted: 
 1. The "simple circle" simulation (simpcirc_gen), where the total habitat area, as well as number and clumping of circular habitat patches can be set
 2. The "random cluster" simulation (randclust_gen), where Reconnect offers a wrapper function applying the random cluster algorithm developped by [Saura and Martinez-Millan 2000](https://link.springer.com/article/10.1023/A:1008107902848) and implemented in the [NLMR package](https://github.com/ropensci/NLMR/) by [Sciaini et al. 2018](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13076).
 
Additionally, Reconnect offers a way to randomly sample "real-world" landscapes of a given size from an input habitat raster file (reallands_gen).

The following shows examples of how to create a library of simulated landscapes with the above mentioned methods.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#########################################################
### clean space
#rm(list=ls(all=TRUE))
#########################################################
## set directories
parent_dir = dirname(dirname(rstudioapi::getActiveDocumentContext()$path))
parent_dir = dirname(getwd())
## source directory
sourcedir  = paste0(parent_dir,"/data")
if(!dir.exists(sourcedir)){dir.create(sourcedir)}
## results directory
resdir     = sourcedir
## folder for results outputs
respath    = paste0(resdir,"/NLMC_demo")
if(!dir.exists(respath)){dir.create(respath)}

##########################################################
## libraries
library(RandomFields)     # NLMR package is relying on this
library(knitr)
library(rmarkdown)
library(sf)               # new spatial vector processing package
library(raster)           # old (but still useful) spatial raster processing
library(stars)            # spatial processing (conversion between sf and raster possible, spatiotemporal datacube analyses)
library(terra)
library(foreach)
library(Reconnect)
## some colors...
JOcols  = c("cyan3","olivedrab3","olivedrab4")
JOcolsr = colorRampPalette(c("navy","lightblue","grey","mediumorchid","cyan3","cyan4","tan","olivedrab3","olivedrab2","olivedrab4"))
JOcolsr1 = colorRampPalette(c("olivedrab3","orange","mediumorchid"))
```


## 1. Simple circle simulation

```{r simpcirc_example}

######################################################################
## Simple circle simulation
######################################################################

######################################################################
### Preparations
## set seed for reproducibility
set.seed(15)
### create a name of the simulation
simname = "SimpCircTest"
## initialize parameters
ncs     = 50                                        ## define number of cells of landscape # 500 cells for dimx and dimy is reasonably good!!
hfrs    = c(0.30)                                   ## define fraction of habitat in landscape, in combination with npatches, this defines patch area
npatches= c(3,10,40)                                ## define number of patches (maximum nr should not be more than 10 times the number of cells..) in combination with hfrs, this defines patch area
sdpas   = 0                                         ## define sd among patch areas
cfs     = 1                                         ## clumping factor: 1 = non-overlapping patches, >1: clumped & overlapping patches
reps    = 1                                         ## number of replicates per run
remedge = TRUE                                      ## remove the edge of the landscape? - i.e. patches cannot go partially outside landscape

######################################################################
### Generate landscape library
simpcirc_gen(simname=simname,ncs=ncs,hfrs=hfrs,npatches=npatches,sdpas=sdpas,cfs=cfs,reps=reps,respath=respath,remedge=remedge) 

######################################################################
### Plot results
files = list.files(paste0(respath,"/",simname))
print(files)
tifs = files[grep(".tif$",files)]
for(tt in tifs) {
 tres = terra::rast(paste0(paste0(respath,"/",simname),"/",tt))
 plot(tres,main=tt)
}

```


## 2. Random Cluster simulation

```{r randomcluster}

######################################################################
## Random cluster simulation
######################################################################

######################################################################
### Preparations
## set seed for reproducibility
set.seed(150)
### create a name of the simulation
simname = "RandClustTest"
## initialize parameters
ncs     = 50                  ## define number of cells of landscape # 500 cells for dimx and dimy is reasonably good!!
hfrs    = c(3/10)             ## define fraction of habitat in landscape, Ai in Saura et al..
cfs     = c(c(1,4,6)/10)      ## clumping factor: pi in [Saura and Martinez-Millan 2000](https://link.springer.com/article/10.1023/A:1008107902848): Note that it is actually not necessary to move pi beyond ca. 0.593 (4-neighborhood clumping rule).
reps    = 1                   ## number of replicates per run
ncores  = 2                   ## number of cores to use for parallel processing

######################################################################
### Generate landscape library
### add missing one: 4.4.2023  RandClust540_nc_0.50_hfr_0.10_cf_1_rep ??
randclust_gen(simname=simname,ncs=ncs,hfrs=hfrs,cfs=cfs,respath=respath,reps=reps,ncores=ncores)

######################################################################
### Plot results
files = list.files(paste0(respath,"/",simname))
print(files)
tifs = files[grep(".tif$",files)]
for(tt in tifs) {
 tres = terra::rast(paste0(paste0(respath,"/",simname),"/",tt))
 plot(tres,main=tt)
}

```


## 3. Real-world samples

```{r realworld_samples}

######################################################################
## Select a number of sample landscapes from a given input raster
######################################################################

######################################################################
### Preparations
## set seed for reproducibility
set.seed(15)

### create a name of the simulation
simname = "RealLandsTest"
## initialize parameters
ncs       = 20              ## define number of cells of landscape # 500 cells for dimx and dimy is reasonably good!!
samplesize = 3              ## number of desired sample landscapes
overlap    = 0.5            ## overlap among sampled landscapes 1=no overlap,if <1 overlap will be bigger, if >1 distance among sampled landscapes will be larger
reps    = 1                 ## number of replicates per run
ncores  = 1                 ## number of cores to use for parallel processing
rast    = as(tres,"Raster") ## land cover raster file containing
names(rast) = "habitat"
### example real-world landscape, from which samples should be taken:
raster::plot(rast,main=names(rast))
######################################################################
### Generate landscape library
reallands_gen(simname=simname,rast=rast,ncs=ncs,samplesize=samplesize,overlap=overlap,respath=respath,habnames=NULL,lcdf=NULL,lccol="hab")

######################################################################
### Plot results
files = list.files(paste0(respath,"/",simname))
print(files)
tifs = files[grep(".tif$",files)]
for(tt in tifs) {
 tres = terra::rast(paste0(paste0(respath,"/",simname),"/",tt))
 plot(tres,main=tt)
}


```


